---
title: "ST 558 Project 3"
author: "Rachel Hencher and Yi Ren"
date: "2022-11-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r, message=FALSE}
library(readr)
library(dplyr)
library(caret)
library(ggplot2)
```

# Data
## Read in and subset data
```{r}
OnlineNewsPopularity <- read_csv("OnlineNewsPopularity.csv") 
OnlineNewsPopularity$url <- NULL
channel <- function(x){
  base <- "data_channel_is_"
  data <- paste0(base,x) %>%  
          noquote()
  news <- OnlineNewsPopularity %>% 
          filter(get(data) == 1) %>% 
          select("Number_Title_Words" = "n_tokens_title",
                 "Number_Content_Words" = "n_tokens_content",
                 "Number_Images" = "num_imgs",
                 "Number_Videos" = "num_videos",
                 "Published_Mon" = ends_with("monday"),
                 "Published_Tues" = ends_with("tuesday"),
                 "Published_Weds" = ends_with("wednesday"),
                 "Published_Thurs" = ends_with("thursday"),
                 "Published_Fri" = ends_with("friday"),
                 "Published_Sat" = ends_with("saturday"),
                 "Published_Sun" = ends_with("sunday"),
                 "Positive_Word_Rate" = "global_rate_positive_words",
                 "Negative_Word_Rate" = "global_rate_negative_words",
                 "Title_Polarity" = "title_sentiment_polarity",
                 "Shares" = "shares")

return(news)
}
```

## Choose an option for the `channel` function argument:  
- *lifestyle*: Is the desired data channel lifestyle?  
- *entertainment*: Is the desired data channel entertainment?  
- *bus*: Is the desired data channel business?  
- *socmed*: Is the desired data channel social media?
- *tech*: Is the desired data channel technology?  
- *world*: Is the desired data channel world?
```{r}
news_data <- channel("lifestyle")
news_data
```

## Split data into train and test
```{r}
set.seed(216)
intrain <- createDataPartition(news_data$Shares, p = 0.7, list = FALSE)
training <- news_data[intrain,]
testing <- news_data[-intrain,]
```

# Summarization
```{r}
# Title length & polarity vs shares
ggplot(training, aes(x = Number_Title_Words, y=Shares)) + 
  geom_point(aes(color = Title_Polarity))
```

# Modeling
## Random forest model
```{r}
rf_model <- train(Shares ~ ., 
                  data = training, 
                  method = "rf", 
                  preProcess = c("center", "scale"), 
                  trControl = trainControl(method = "cv", number = 5), 
                  tuneGrid = expand.grid(mtry = c(1:14)))
```
                         
## Boosted tree model
```{r, results="hide"}
tunegrid <- expand.grid(interaction.depth = 1:4,
                        n.trees = c(25, 50, 100, 150, 200),
                        shrinkage = 0.1,
                        n.minobsinnode = 10)

gbm_model <- train(Shares ~.,
                 data = training,
                 method = "gbm",
                 preProcess = c("center", "scale"),
                 trControl = trainControl(method = "cv", number = 5),
                 tuneGrid = tunegrid)
```

# Comparison
## Apply model for prediction
```{r}
rf_predict <- predict(rf_model, newdata = testing)
```

```{r}
gbm_predict <- predict(gbm_model, newdata = testing)
```


## Model performance
```{r, eval=FALSE}
confusionMatrix(data = testing$Shares, reference = rf_predict)
```

```{r, eval=FALSE}
confusionMatrix(data = testing$Shares, reference = gbm_predict)
```


# Automation