---
title: "Project3"
author: "Rachel Hencher and Yi Ren"
date: "2022-11-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r, message=FALSE}
library(readr)
library(dplyr)
library(caret)
```

# Data
## Read in and subset data
```{r}
OnlineNewsPopularity <- read_csv("OnlineNewsPopularity.csv") 
OnlineNewsPopularity$url <- NULL
channel <- function(x){
  base <- "data_channel_is_"
  data <- paste0(base,x) %>% 
          noquote()
  news <- OnlineNewsPopularity %>% 
          filter(get(data) == 1) %>% 
          select("Number_Title_Words" = "n_tokens_title",
                 "Number_Content_Words" = "n_tokens_content",
                 "Number_Images" = "num_imgs",
                 "Number_Videos" = "num_videos",
                 starts_with("weekday"),
                 "Positive_Word_Rate" = "global_rate_positive_words",
                 "Negative_Word_Rate" = "global_rate_negative_words",
                 "Title_Polarity" = "title_sentiment_polarity",
                 "Shares" = "shares")
  return(news)
}
```

## Set up options for the `channel` function argument:  
- *lifestyle*: Is the desired data channel lifestyle?  
- *entertainment*: Is the desired data channel entertainment?  
- *bus*: Is the desired data channel business?  
- *socmed*: Is the desired data channel social media?
- *tech*: Is the desired data channel technology?  
- *world*: Is the desired data channel world?
```{r}
lifestyle <- channel("lifestyle")
entertainment <- channel("entertainment")
bus <- channel("bus")
socmed <- channel("socmed")
tech <- channel("tech")
world <- channel("world")

news_data <- channel("world")
news_data
```

## Split data into train and test
```{r}
set.seed(216)
intrain <- createDataPartition(data$Shares, p = 0.7, list = FALSE)
lifestyletrain <- data[intrain,]
lifestyletest <- data[-intrain,]
```

# Summarization

# Modeling
## Boosted tree model
```{r, results='hide'}
tunegrid <- expand.grid(interaction.depth = 1:4,
                        n.trees = c(25, 50, 100, 150, 200),
                        shrinkage = 0.1,
                        n.minobsinnode = 10)

gbmTune <- train(Shares ~.,
                 data = lifestyletrain,
                 method = "gbm",
                 trControl = trainControl(method = "cv",
                                          number = 5),
                 tuneGrid = tunegrid)
```


# Comparison
## Apply model for prediction
pred_gbm <- predict(gbmTune, newdata = lifestyletest)

## Model performance
confusionMatrix(data = lifestyletest$shares, reference = pred_gbm)

# Automation