---
title: "ST 558 Project 3"
author: "Rachel Hencher and Yi Ren"
date: "2022-11-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packages
```{r packages, message=FALSE}
library(readr)
library(dplyr)
library(caret)
library(ggplot2)
```

# Data
## Read in and subset data
```{r subset, message=FALSE}
OnlineNewsPopularity <- read_csv("OnlineNewsPopularity.csv") 
OnlineNewsPopularity$url <- NULL
channel <- function(x){
  base <- "data_channel_is_"
  data <- paste0(base,x) %>%  
          noquote()
  news <- OnlineNewsPopularity %>% 
          filter(get(data) == 1) %>% 
          select("Number_Title_Words" = "n_tokens_title",
                 "Number_Content_Words" = "n_tokens_content",
                 "Number_Images" = "num_imgs",
                 "Number_Videos" = "num_videos",
                 starts_with("weekday_is"),
                 "Positive_Word_Rate" = "global_rate_positive_words",
                 "Negative_Word_Rate" = "global_rate_negative_words",
                 "Title_Polarity" = "title_sentiment_polarity",
                 "Shares" = "shares")
  
news$Weekday <- as.factor(ifelse(news$weekday_is_monday == 1, "Mon",
                                 ifelse(news$weekday_is_tuesday == 1, "Tues", 
                                        ifelse(news$weekday_is_wednesday == 1, "Weds", 
                                               ifelse(news$weekday_is_thursday , "Thurs",
                                                      ifelse(news$weekday_is_friday == 1, "Fri",
                                                             ifelse(news$weekday_is_saturday == 1, "Sat", "Sun")))))))

news_final <- news %>%
  select(-starts_with("weekday_is"))
return(news_final)
}
```

## Choose an option for the `channel` function argument:  
- *lifestyle*: Is the desired data channel lifestyle?  
- *entertainment*: Is the desired data channel entertainment?  
- *bus*: Is the desired data channel business?  
- *socmed*: Is the desired data channel social media?
- *tech*: Is the desired data channel technology?  
- *world*: Is the desired data channel world?

```{r channel}
news_data <- channel("lifestyle")
news_data
```

## Split data into train and test
```{r split}
set.seed(216)
intrain <- createDataPartition(news_data$Shares, p = 0.7, list = FALSE)
training <- news_data[intrain,]
testing <- news_data[-intrain,]
```

# Summarization
```{r summary}
# Summary stats for all variables
summary(training)
# Boxplot of weekday vs shares
ggplot(training, aes(x = Weekday, y = Shares)) +
  geom_boxplot(color = "royal blue") +
  scale_y_continuous(trans="log10")
# Scatterplot of title length & polarity vs shares
ggplot(training, aes(x = Number_Title_Words, y = Shares)) + 
  geom_point(aes(color = Title_Polarity))
```

# Modeling
## Set up cross validation
```{r control}
control <- trainControl(method = "cv", number = 5)
```

## Linear regression models
```{r lasso}
lasso_model <- train(Shares ~ .,
                   data = training,
                   method ='lasso',
                   preProcess = c("center", "scale"),
                   trControl = control)
predict(lasso_model$finalModel, type = "coef")
lasso_model$bestTune
```

```{r fwstep}
fwdstep_model <- train(Shares ~ .,
                   data = training,
                   method ='glmStepAIC',
                   preProcess = c("center", "scale"),
                   trControl = control,
                   direction = "forward",
                   trace = FALSE)
fwdstep_model
```

## Random forest model
```{r rf}
rf_model <- train(Shares ~ ., 
                  data = training, 
                  method = "rf", 
                  preProcess = c("center", "scale"), 
                  trControl = control, 
                  tuneGrid = expand.grid(mtry = (1:ncol(training))/3))
rf_model
```
                         
## Boosted tree model
```{r gbm}
gbm_model <- train(Shares ~ .,
                     data = training,
                     method = "gbm",
                     trControl = control,
                     preProcess = c("center", "scale"),
                     verbose = FALSE)
gbm_model
```

# Comparison
## Apply model for prediction
```{r prediction}
lasso_predict <- predict(lasso_model, newdata = testing)
fwdstep_predict <- predict(fwdstep_model, newdata = testing)
rf_predict <- predict(rf_model, newdata = testing)
gbm_predict <- predict(gbm_model, newdata = testing)
```

## Model performance
```{r performance}
Lasso <- postResample(lasso_predict, obs = testing$Shares)
Forward_Stepwise <- postResample(fwdstep_predict, obs = testing$Shares)
Random_Forest <- postResample(rf_predict, obs = testing$Shares)
Boosted_Tree <- postResample(gbm_predict, obs = testing$Shares)

rbind(Lasso, Forward_Stepwise, Random_Forest, Boosted_Tree)
```


# Automation
